---
title: "Exponential Distribution & Application of Central Limit Theorem"
author: "Nikhil Gupta"
date: "June 25, 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This report will investigate the exponential distribution in R and compare it with the Central Limit Theorem. Exponential distribution with lambda=0.2 will be used for the purpose with simulation count of 1000.To validate the central limit theorem, average of 40 exponentials wil be taken.


##Exponential random variable simulation 

Firstly, random variable with exponential distribution has been simulated (based on theoretical & sample central measures). This has been used to calculate the required confidence intervals.
```{r echo=TRUE}
lambda =0.2
par(mfrow=c(1,2)) 
exp1= rexp(1000,lambda)
hist(exp1, col = "blue", ylab = "Frequency")
theoretical_mean = 1/lambda 
sample_mean = mean(exp1) 
sample_sd=sd(exp1)
theoretical_sd=1/lambda
abline(v=sample_mean, col="green", lwd=6, lty=2);
abline(v=theoretical_mean, col="red", lwd=3);
legend("topright", c("Sample_mean", "Theoretical_mean"), fill=c("green", "red"),cex=0.7)

cat("Sample variance of exponential distribution:", sample_sd^2,"\n")
cat("Sample two-sided 95% confidence interval:",c((sample_mean-1.96*sample_sd),(sample_mean+1.96*sample_sd)),"\n")
cat("Population variance of exponential distribution:", theoretical_sd^2,"\n")
cat ("Population two-sided 95% confidence interval",c((theoretical_mean-1.96*theoretical_sd),(theoretical_mean+1.96*theoretical_sd)),"\n")
cat("\n")
summary(exp1)
cat("\n")
```

##Exponential random variables mean simulation 

Secondly, mean of 40 random variables with exponential distribution has been simulated (based on theoretical & sample central measures). This should tend towards normal random variable as per central limit theorem. This has been validated & required confidence intervals have been calculated.

```{r}

exp_avg =vector(mode="numeric", length=1000)
 for (i in 1:1000)
 { exp_avg[i]= mean(rexp(40,lambda))
 }
 
 hist(exp_avg, col = "blue", ylab = "Frequency")
 theoretical_mean1 = 1/lambda 
 sample_mean1 = mean(exp_avg) 
 sample_sd1=sd(exp_avg)
 theoretical_sd1=1/(lambda* (40^0.5))
 abline(v=sample_mean1, col="green", lwd=6, lty=2);
 abline(v=theoretical_mean1, col="red", lwd=3);
 legend("topright", c("Sample_mean", "Theoretical_mean"), fill=c("green", "red"),cex=0.6)
 
cat("Sample variance of averaged exponential distribution:", sample_sd1^2,"\n")
cat("Sample two-sided 95% confidence interval:",c((sample_mean1-1.96*sample_sd1),(sample_mean1+1.96*sample_sd1)),"\n")
cat("Population variance of averaged exponential distribution:", theoretical_sd1^2,"\n")
cat ("Population two-sided 95% confidence interval",c((theoretical_mean1-1.96*theoretical_sd1),(theoretical_mean1+1.96*theoretical_sd1)),"\n")
cat("\n")
summary(exp_avg) 
 
```
